# Reproducible Research Assignment 1

The data consists of two month of data from a personal activity monitoring device like fitbit or jawbone collected during the months of October and November 2012. 

## Loading and preprocessing the data
Load data

```{r}
data <- read.csv(file="activity.csv",header=T);
str(data);
```

## What is mean total number of steps taken per day?
The total steps taken by day calculated by aggregating day (factor),

```{r}
daily_steps <- aggregate(data$steps, by=list(data$date), FUN=sum);
daily_total_steps <- daily_steps[!(is.na(daily_steps$x)),2];
hist(daily_total_steps,xlab="Daily Total Steps");
```

The mean and median is calculated for the total steps.

```{r}
mean(daily_steps$x,na.rm=T);
median(daily_steps$x,na.rm=T);
```

## What is the average daily activity pattern?
The average steps in 5 minutes interval is calculated and ploted.

```{r}
interval_means <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=TRUE);
plot(interval_means, type = "l", xlab="minutes", ylab="number of steps");
```
The interval with the maximum steps over all days is calculated.
```{r}
max_mean <- interval_means[interval_means$x == max(interval_means$x),];
max_mean;
```

## Missing data
There are missing data in the data set.

The numbers of meassurments with NA for steps are:
```{r}
sum(is.na(data$steps));
```
Those meassurments with NA will be replaced by the mean of the coresponding interval (which has been calculated in the above step).

```{r}
merged <- merge(x=data, y=interval_means, by.x="interval", by.y="Group.1");
my.na <- is.na(merged$steps);
merged[my.na, 2] <- merged[my.na, 4];
data2 <- data.frame(steps=merged$steps, date=merged$date, interval=merged$interval)

interval_means <- aggregate(data2$steps, by=list(data2$interval), FUN=mean);
plot(interval_means$x, type = "l", xlab = "minutes", ylab = "number of steps");
```
The mean and median is calculated for the total steps of the new data set.

```{r}
daily_steps <- aggregate(data2$steps, by=list(data2$date), FUN=sum);
mean(daily_steps$x,na.rm=T);
median(daily_steps$x,na.rm=T);
```
With that replacement only the median has changed for just one step.

## Weekday and Weekend
Creating a new column with the factor weekday and weekend.

```{r}
library(timeDate);
data2 = within(data2, {
    daytype = as.factor(ifelse(isWeekday(data2$date), "weekday", "weekend"))
 })
str(data2);
```

Than two plots are created to show the average over weekdays and weekends.
```{r}
par(mfrow = c(2,1));
X <- split(data2, data2$daytype);
weekday_interval_means <- aggregate(X[[1]]$steps, by=list(X[[1]]$interval), FUN=mean);
weekend_interval_means <- aggregate(X[[2]]$steps, by=list(X[[2]]$interval), FUN=mean);
plot(weekday_interval_means$x, type = "l", xlab = "minutes", ylab = "number of steps", main = "Weekday");
plot(weekend_interval_means$x, type = "l", xlab = "minutes", ylab = "number of steps", main = "Weekend");
```